Approximation as Differentiation: Notes

### 2025-06-07 main/approx-as-tangents.tex
Unresolved citations.

Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory. Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry}. We now introduce these concepts and how they relate to \GPS.

@@

Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality. This was not successful (see the $\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS. A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12). We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.

>>>>

Our thesis is that \GPS is a form of differentiable programming, where tangents are not linear approximations of functions but instead are qualitiative approximations of elements in the sense of domain theory. Smooth functions in this setting are Berry's \emph{stable functions} \cite{berry79,berry82}. We now introduce these concepts and how they relate to \GPS.

@@

Stability is a property that can be required of monotone functions that was invented by Berry~\cite{berry79} in an attempt to capture sequentiality. This was not successful (see the $\mathrm{gustave}$ function in \autoref{ex:parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS. A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12). We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.
